Design and Implementation of Probabilistic Programming Language Anglican

### 2017-02-13 paper/paper.tex
typo fixed: this -> these
in the middle of revising the intro slightly.

Anglican is implemented in just this way --- a macro facility
provided by Clojure, a Lisp dialect, is used both to extend
Clojure with constructs that delimit probabilistic code, and to
alter the operational semantics of Clojure expressions inside
probabilistic code fragments. Anglican claims its right to count
as a separate language because of the ubiquitous probabilistic
execution semantics rather than a different syntax,
which is actually an advantage rather than a drawback ---
Clojure programmers only need to know how to specify the
boundaries of Anglican programs, but can use familiar Clojure
syntax to write probabilistic code.

@@

An execution of the query may result in the following
sequence of checkpoint invocations:
\begin{lstlisting}[style=default]
  (sample 'C1 ...)
  (sample 'C2 ...)
  (sample 'C2 ...)
  (sample 'C1 ...)
  (sample 'C1 ...)
  (sample 'C1 ...)
  (sample 'C2 ...)
  (sample 'C3 ...)
\end{lstlisting}
According to the addressing scheme, the addresses generated
for this invocations are
\begin{lstlisting}[style=default]
  [C1 0][C2 0][C2 1][C1 16][C1 17][C1 18][C2 16][C3 0]
\end{lstlisting}
If the program is run by a Metropolis-Hastings algorithm, then a
small change usually takes place in the sequence of checkpoints
with each invocation, and the new sequence may become
\begin{lstlisting}[style=default]
  (sample 'C1 ...)
  (sample 'C2 ...)
  (sample 'C1 ...)
  (sample 'C1 ...)
  (sample 'C2 ...)
  (sample 'C2 ...)
  (sample 'C3 ...)
\end{lstlisting}
The addresses for the new sequence are
\begin{lstlisting}[style=default]
  [C1 0][C2 0][C1 16][C1 17][C2 16][C2 17][C3 0]
\end{lstlisting}
Despite the change, the correspondence between checkpoints of
similar types occurring in similar positions (relative positions
in contiguous subsequences of a certain type) is preserved, and
drawn values can be reused efficiently:

>>>>

Anglican is implemented in just this way --- a macro facility
provided by Clojure, a Lisp dialect, is used both to extend
Clojure with constructs that delimit probabilistic code, and to
alter the operational semantics of Clojure expressions inside
probabilistic code fragments. Anglican claims its right to count
as a separate language, rather than being a Clojure library, 
because of the ubiquitous probabilistic
execution semantics rather than a different syntax,
which is actually an advantage rather than a drawback ---
Clojure programmers only need to know how to specify the
boundaries of Anglican programs, but can use familiar Clojure
syntax to write probabilistic code.

\hy{I am here.}

Anglican programs are executed by inference algorithms, which
answer probabilistic queries on these programs approximately 
often using a large number of samples. As already mentioned,
this execution is significantly different from that of 
usual programs. An Anglican program is executed multiple times, often hundreds
of thousands or even millions of times for a single inference
task. Random choices may affect which parts of the program code
are executed and how often. Many inference algorithms require
re-running the program multiple times partially, from a certain
point on.  Different executions may employ different random
choices. However, for efficient inference a correspondence
between random choices in different executions should be
maintained. These are just some of the challenges which were
faced and solved during development of Anglican.

@@

An execution of the query may result in the following
sequence of checkpoint invocations:
\begin{lstlisting}[style=default]
  (sample 'C1 ...)
  (sample 'C2 ...)
  (sample 'C2 ...)
  (sample 'C1 ...)
  (sample 'C1 ...)
  (sample 'C1 ...)
  (sample 'C2 ...)
  (sample 'C3 ...)
\end{lstlisting}
According to the addressing scheme, the addresses generated
for these invocations are
\begin{lstlisting}[style=default]
  [C1 0][C2 0][C2 1][C1 16][C1 17][C1 18][C2 16][C3 0]
\end{lstlisting}
If the program is run by a Metropolis-Hastings algorithm, then a
small change usually takes place in the sequence of checkpoints
with each invocation, and the new sequence may become
\begin{lstlisting}[style=default]
  (sample 'C1 ...)
  (sample 'C2 ...)
  (sample 'C1 ...)
  (sample 'C1 ...)
  (sample 'C2 ...)
  (sample 'C2 ...)
  (sample 'C3 ...)
\end{lstlisting}
The addresses for the new sequence are
\begin{lstlisting}[style=default]
  [C1 0][C2 0][C1 16][C1 17][C2 16][C2 17][C3 0]
\end{lstlisting}
Despite the change, the correspondence between checkpoints of
similar types occurring in similar positions (relative positions
in contiguous subsequences of a certain type) is preserved, and
drawn values can be reused efficiently:
